---
title: "Maps"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')


library(dplyr)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(dslabs)
library(datasauRus)
```


## Your Turn 1

The artificial dataset used in this exercise can be loaded as follows:

```{r}
mountain <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/mountain.csv")
```


### (a) Apply `theme_light()` to the histogram.

```{r}
mountain_plot <- ggplot(mountain) +
  geom_histogram(
    aes(x = Height, fill = Location), 
    binwidth = 1,
    color = "white"
  ) +
  scale_fill_manual(values = c("green3", "lightblue3"))

# Now, apply the light theme
mountain_plot + _________

```


### (b) Close "gap" between bars and axis

```{r}
# No space below the bars but 10% above them
last_plot() + 
  scale_y_continuous(expand = expansion(mult = c(0, .1)))
```

### (c) Remove legend title

```{r}
last_plot() +
  labs(fill = NULL)
```

### (d) Let's remove border and minor grid lines

```{r}
last_plot() +
  theme(
    panel.border = element_blank(),
    panel.grid.minor = element_blank()
  )
```

### (e) Reposition legend & remove the background

```{r}
last_plot() +
  theme(
    legend.position = c(.15, .85),
    legend.background = element_blank()
  )
```

### (f) Extra: Change the axes labels using `labs()`

```{r}
# your code

last_plot() +
  labs(x = "___________", y = "_________")

```



## Your Turn 2

In today's class we created cloropleth maps of states in the US based on ACS data. 

```{r}
states <- map_data("state")
ACS <- ACS <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/ACS.csv")
ACS <- dplyr::filter(ACS, !(region  %in% c("Alaska", "Hawaii"))) # only 48+D.C.
ACS$region <- tolower(ACS$region)  # lower case (match states regions)
```

### (a) Mapping median income

Create a cloropleth plot that uses color to create a `MedianIncome`  map of the US.

```{r}
# map median income
ggplot(data=ACS) + coord_map() + 
  geom_map(aes(map_id = ______, fill = ___________), map = states) +
  expand_limits(x=states$long, y=states$lat) + ggtitle("Median Income")
```

### (b) Mapping deviations from national median income

The median income in the US in 2016 was estimated to be $27,000. Redraw your map in (a) to visualize each state's deviation from national median income. 

```{r}
# compare state income to national income

ggplot(data=ACS) + coord_map() + 
  geom_map(aes(map_id = region, fill = _____________), map = states) +
  expand_limits(x=states$long, y=states$lat) + ggtitle("Deviation from national median income")
```

### (c) Changing numerically scaled color

You should use a *diverging* color for (b) to highlight larger deviations from the national median. Add `scale_fill_distiller(type = "div")` to the map from (b) and select a diverging palette.   

```{r}
# change to a diverging color
ggplot(data=ACS) + coord_map() + 
  geom_map(aes(map_id = region, fill = MedianIncome - 27000), map = states) +
  expand_limits(x=states$long, y=states$lat) + ggtitle("Deviation from national median income") + 
  __________________________
```

### (d) Fixing a midpoint on a diverging scale

Use `scale_fill_gradient2` to fix a midpoint scale value at a white color, with diverging colors for larger positive and negative values. Apply this color to your map in (b) and fix the `midpoint` at an appropriate value. 

```{r}
# change to a gradient fill color
ggplot(data=ACS) + coord_map() + 
  geom_map(aes(map_id = region, fill = MedianIncome - 27000), map = states) +
  expand_limits(x=states$long, y=states$lat) + ggtitle("Deviation from national median income") + 
 _________________________________
```

